---
title: "agds_report_01"
author: "Naomi Huber"
date: "2023-05-15"
output: html_document
---

## Dimensions of a circle
```{r}
r <- 4

a <- as.character((r**2)*pi)
circumference <- as.character(2*r*(pi**2))

print(c(a, circumference))
```
My code defines that r saves the number 4. Afterwards, I save in the variable a the code as.character((r**2)*pi) which is basically the calculation formula for the calculation of a circle area. I save also the calculation of the circumference in the variable circumference. Then i print the vector a (which saves the calculated area) and the circumference of this circle. 

## Sequence of numbers
```{r}
x <- c(seq(from = 0, to = pi, length.out = 5))

cat("A sequence of numbers from 0 to Pi: ", x, sep = " / ")
```
In the calculation above I saved a sequence of numbers from 0 to Pi in the variable x. Then I printed a sentence with the five possible solutions which have all the same difference.

## Gauss sum
```{r}
x <- c(seq(from = 1, to = 100))
y <- c(seq(from = 100, to = 1))

new_vector <- x+y
fifty <- length(x+y)/2

cat("The Gauss sum is: ", new_vector[1]*fifty)
```
In this code I defined two variables (x and y) which save two sequences of numbers each. The variable x saves all the numbers from 1 to 100 and y saves all the numbers (downwards) from 100 to 1. Then I calculate the Gauss sum with new_vector and fifty (variables). 

## Magic trick algorithm
```{r}
x <- as.integer(5)
x1 <- x
x <- x + 1
x <- x * 2
x <- x1 + 4
x <- x1/2
x - x1
```
I overwrote x and changed it's value.


## Vectors
```{r}
datasets::rivers
class <- class(datasets::rivers)
length <- length(datasets::rivers)
mean <- mean(datasets::rivers)
median <- median(datasets::rivers)
minimum <- min(datasets::rivers)
maximum <- max(datasets::rivers)
```

I opened the object dataset::rivers and defined some important things in this object (like the class, length, mean, median, minimum, maximum).
